---
title: "Homework 6"
output: html_document
date: "2023-12-21"
---


```{r}
library(tidyverse)
library(viridis)
library(modelr)
library(ggplot2)
library(purrr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


### Problem 1 - Solution Already Provided

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2

Importing and reading in the data:

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Build a regression model for `tmax`

```{r}
model_tmax <- lm(tmax ~ tmin + prcp, data = weather_df)
summary(model_tmax)
```

Bootstrap analysis

```{r}
set.seed(123)
```

Writing a function to generate 5000 bootstraps samples

```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}
```

5000 bootstraps samples:

```{r}
boot_straps = 
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df))
  )

boot_straps
```

Combine all the bootstraps results

```{r}
bootstrap_results <- 
  boot_straps |> 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin + prcp, data = .x) ),
    results = map(models, broom::glance)) |>
  select(-strap_sample, -models) |>
  unnest(results) 
```

Compute the R-Squared

```{r}
r_squared <- 
  bootstrap_results |>
  select(r.squared, strap_number)
```

Generating estimates

```{r warning=FALSE}
bootstrap_results2 = 
  boot_straps |> 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin + prcp, data = .x) ),
    results = map(models, broom::tidy)) |> 
  select(-strap_sample, -models) |> 
  unnest(results) 

```

Log of the product of beta estimates

```{r}
log_betas <-  
  bootstrap_results2 |>
  group_by(strap_number) |>
  summarise(log_betas = log(estimate[2] * estimate[3])) |>
  select(log_betas, strap_number)
```

Density Plot of R-squared

```{r}
r_squared_sd <-
  r_squared |>
  summarise(r_squared_sd = sd(r.squared)) |>
  pull(r_squared_sd)

r_squared_mean <-
  r_squared |>
  summarise(r_squared_mean = mean(r.squared)) |>
  pull(r_squared_mean)

r_squared |>
  ggplot(aes(x = r.squared)) + geom_density()+
  labs(title = "Distribution of R-squared Estimates")
```

The density curve of R-Squared $\hat r^2$ is skewed to the left. The mean is `r round(r_squared_mean,4)` and standard deviation is`r round(r_squared_sd,4)`.


Density plot of the product of beta estimates

```{r}
  beta_sd <- 
  log_betas |>
  summarise(beta_sd = sd(as.numeric(log_betas),na.rm = TRUE)) |>
  pull(beta_sd)

  betas_mean <- 
  log_betas |>
  summarise(betas_mean = mean(as.numeric(log_betas), na.rm = TRUE)) |>
  pull(betas_mean)

log_betas |>
  ggplot(aes(x = log_betas)) + geom_density() +
  labs(title = "Distribution of Beta Estimates")
```

The density curve of the the product of two beta estimates is left-skewed. The mean is `r round(betas_mean,4)`, and the standard deviation is `r round(beta_sd,4)`.


95% CI of the R-Squared
```{r}
CI_interval <-
  r_squared |>
  summarize(ci_lower = quantile(r.squared, 0.025),
            ci_upper = quantile(r.squared, 0.975)) 

CI_result_lower <- CI_interval |> pull(ci_lower)
CI_result_upper <- CI_interval |> pull(ci_upper)
```

The 95% Confidence Interval of the R-squared estimates $\hat r^2$ is (`r round(CI_result_lower,4)`,`r round(CI_result_upper,4)`).


95% CI of log of the product of beta estimates
```{r}
CI_interval2 <-
  log_betas |>
  summarize(ci_lower = quantile(log_betas, 0.025, na.rm = TRUE),
            ci_upper = quantile(log_betas, 0.975, na.rm = TRUE))

CI_result_lower2 <- CI_interval2 |> pull(ci_lower)
CI_result_upper2 <- CI_interval2 |> pull(ci_upper)
```

The 95% Confidence Interval of the product of beta estimates $\log(\hat\beta_1\times \hat\beta_2)$ is (`r round(CI_result_lower2,4)`,`r round(CI_result_upper2,4)`).


### Problem 3
















